---
title: Integration of multiple experiments for the ccRCC project
author: 
- name: Nick Borcherding
  email: ncborch@gmail.com
  affiliation: Washington University in St. Louis, School of Medicine, St. Louis, MO, USA
date: "June 25, 2020"
output:
  BiocStyle::html_document:
    toc_float: true

---

```{r, echo=FALSE, results="hide", message=FALSE}
knitr::opts_chunk$set(error=FALSE, message=FALSE, warning=FALSE)
library(BiocStyle)
```


# Preperation

## Loading Libaries

In general I like to load libraries here that we will use universally, and then call other libraries when we need them in the code chunks that are relevant. 

```{r}
suppressPackageStartupMessages(library(Seurat))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
```

## Loading Data

```{r}
integrated <- readRDS("./data/Processed/integrated_Cluster.rds")
load("./data/Processed/completeMeta.rda")
```

## Selecting Color Palette

```{r setup, include=FALSE}
colorblind_vector <- colorRampPalette(c("#FF4B20", "#FFB433", "#C6FDEC", "#7AC5FF", "#0348A6"))
```

***

# Initial Subsetting and Analysis

We will use the **Minor** variable in the integrated seurat object meta data that we defined in the integration step, filtering for the "CD5.T.Cell" and "Tregs". 

```{r}
Tcells <- subset(integrated, Minor == "CD4.T.Cell" | Minor ==  "Treg")
Tcells <- NormalizeData(Tcells)
rm(integrated)
```

We are also going to create a subdirectory for all T cell analyses:
```{r}
dir.create("./DataAnalysis/CellType/CD4.Tcells")
```




## Reclustering 

Isolating just the T cells, stripping the seurat object of non-"RNA" values using *DietSeurat()* and then spliting the object to create a list that can be fed into **SCTransform()**
```{r eval=FALSE}
Tcells@meta.data$Run <- paste0(Tcells@meta.data$sample, "_", Tcells@meta.data$type) #Define the sequence run
Tcells[["percent.mt"]] <- PercentageFeatureSet(Tcells, pattern = "^MT-") #Calculate %mito

Tcells <- DietSeurat(Tcells, assays = "RNA")
T.list <- SplitObject(object = Tcells , split.by = "Run")

for (i in 1:length(T.list)) {
    T.list[[i]] <- SCTransform(T.list[[i]], verbose = FALSE, conserve.memory=TRUE)
}
rm(Tcells)
```

Integrating the data as above with the whole data set.
```{r eval = FALSE}
options(future.globals.maxSize= 2621440000)
features <- SelectIntegrationFeatures(object.list = T.list, nfeatures = 3000)
T.list <- PrepSCTIntegration(object.list = T.list, anchor.features = features,
verbose = FALSE)
T.anchors <- FindIntegrationAnchors(object.list = T.list, normalization.method = "SCT",
anchor.features = features, verbose = FALSE, k.filter = 40)
Tcells <- IntegrateData(anchorset = T.anchors, normalization.method = "SCT", verbose = FALSE)
rm(T.list)
rm(T.anchors)
saveRDS(Tcells, file = "./data/Processed/CD4.Tcells_Precluster.rds")
```

Now we can process the intgreated T cells and identify clusters.
```{r eval = FALSE}
Tcells <- readRDS("./data/Processed/CD4.Tcells_Precluster.rds")


dir.create("./DataAnalysis/CellType/CD4.Tcells/DimensionInputs")
for (i in c(10,15,20,25,30,35,40)) {
    
    integrated2 <- ScaleData(object = Tcells, verbose = FALSE)
    integrated2 <- RunPCA(object = integrated2, npcs = 40, verbose = FALSE)
    integrated2 <- RunUMAP(object = integrated2, reduction = "pca", 
        dims = 1:i) #dimensions for  UMAP
    for (x in c(10,15,20,25,30,35,40)) { #Dimensions for neighbors
       integrated2 <- FindNeighbors(object = integrated2, dims = 1:x)
       for(y in c(0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) { #Resolution
           integrated2 <- FindClusters(object = integrated2, resolution = y, force.recalc=T)
           plot1 <- DimPlot(integrated2, reduction = "umap", pt.size = 0.5, group.by = "orig.ident")
           ggsave(path = "./DataAnalysis/CellType/CD4.Tcells/DimensionInputs", file = paste("UMAP", i, "_Neighbors", x, "_resolution", y, "_OrigIdent.pdf", sep=""), plot1, width=4.25, height=3)
           plot2 <- DimPlot(object = integrated2, reduction = 'umap', label = T) + NoLegend()
            ggsave(path = "./DataAnalysis/CellType/CD4.Tcells/DimensionInputs", file = paste("UMAP", i, "_Neighbors", x, "_resolution", y, "_clusters.pdf", sep=""), plot2, width=3.5, height=3)
       }
    }
}
rm(integrated2)

```


```{r eval = FALSE}
Tcells <- ScaleData(object = Tcells, verbose = FALSE)
Tcells <- RunPCA(object = Tcells, npcs = 40, verbose = FALSE)
Tcells <- RunUMAP(object = Tcells, reduction = "pca", 
    dims = 1:35)
Tcells <- FindNeighbors(object = Tcells, dims = 1:40, force.recalc = T)
Tcells <- FindClusters(object = Tcells, resolution = 0.4, force.recalc=T)
Tcells <- subset(Tcells, idents = c(0,1,2,3,4,5,6,7))
```

### Visualizing the new sub-clusters

```{r}
Tcells@meta.data$newCluster <- Tcells@active.ident
DimPlot(object = Tcells, reduction = 'umap', label = T) + NoLegend()
ggsave(path = "./DataAnalysis/CellType/CD4.Tcells", filename="IntegratedObject_byCluster.eps", width=3.5, height=3)
DimPlot(object = Tcells, reduction = 'umap', group.by = "type") 
ggsave(path = "./DataAnalysis/CellType/CD4.Tcells", filename="IntegratedObject_byType.eps", width=3.75, height=3)
```

Examining the relative proportion on the new clusters
```{r}
freq_table <- table(Tcells@active.ident, Tcells$type)

tTotal <- 12239
pTotal <- 21160
kTotal <- 1778
totals <- c(kTotal,pTotal,tTotal)

for (i in 1:3) { 
    freq_table[,i] <- freq_table[,i]/totals[i]
}

freq_table <- reshape2::melt(freq_table)

ggplot(freq_table, aes(x=as.factor(Var1), y=value, fill=Var2)) + 
  geom_bar(stat="identity", position="fill", color="black", lwd=0.25) + 
  theme(axis.title.x = element_blank()) + 
   scale_fill_manual(values=rev(colorblind_vector(3))) + 
  theme_classic() + 
    xlab("Clusters") + 
    coord_flip() 
ggsave(path = "DataAnalysis/CellType/CD4.Tcells/", file = "relativeContribution_byClusterType_scaled.pdf", height=5, width=5)


ggplot(freq_table, aes(x=as.factor(Var1), y=value, fill=Var2)) + 
  geom_bar(stat="identity", color="black", lwd=0.25) + 
  theme(axis.title.x = element_blank()) + 
   scale_fill_manual(values=rev(colorblind_vector(3))) + 
  theme_classic() + 
    xlab("Clusters") + 
    coord_flip()
ggsave(path = "DataAnalysis/CellType/CD4.Tcells/", file = "totalContribution_byClusterType_scaled.pdf", height=5, width=5)

```

## Cell Cycle 

Cell cycle regression as described in the [Satija lab website](https://satijalab.org/seurat/v3.1/cell_cycle_vignette.html).  

```{r}
cc.genes <- Seurat::cc.genes.updated.2019
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
```

Now we can perform cell cycle scoring with the genes. For now, I am not going to regress using the assingments or save the integrated object with the calculations.

```{r}
Tcells <- CellCycleScoring(Tcells, s.features = s.genes, g2m.features = g2m.genes, set.ident = FALSE)
```

Like above with the contribution by condition (**type**), we can now look at the phases assignments by cluster and condition. We do not need to scale if we seperate **type** into seperate bar graphs.

```{r}
dir.create("./DataAnalysis/CellType/CD4.Tcells/CellCycle")
freq_table <- Tcells[[]]
freq_table <- freq_table[,c("type", "Final_clusters", "Phase")]
freq_table <- subset(freq_table, Phase != "Undecided") #removing undecided phases
freq_table <- freq_table %>%
    group_by(type, Final_clusters, Phase) %>%
    summarise(n = n())

freq_table$Phase <- factor(freq_table$Phase, levels = c("G1", "S", "G2M")) #ordering phases
freq_table$Final_clusters <- factor(freq_table$Final_clusters, levels = c(1,2,3,8,12,13,14,17,18,21,23,24))

ggplot(freq_table, aes(x=Final_clusters, y=n, fill=Phase)) + 
  geom_bar(stat="identity", position="fill", color="black", lwd=0.25) + 
  theme(axis.title.x = element_blank())+ 
    facet_grid(type ~.) + 
   scale_fill_manual(values=colorblind_vector(3)) + 
  theme_classic()
ggsave(path = "./DataAnalysis/CellType/CD4.Tcells/CellCycle", file = "CellCycle_byCluster_byType.pdf", height=6, width=6)

ggplot(freq_table, aes(x=Final_clusters, y=n, fill=Phase)) + 
  stat_summary(geom="bar", position="fill", color="black", lwd=0.25) + 
  theme(axis.title.x = element_blank())+ 
   scale_fill_manual(values=colorblind_vector(3)) + 
  theme_classic()
ggsave(path = "./DataAnalysis/CellType/CD4.Tcells/CellCycle", file = "CellCycle_byCluster.pdf", height=2, width=6)
```


```{r}
Tcells$Phase <- factor(Tcells$Phase, levels = c("G1", "S", "G2M")) #ordering phases
DimPlot(object = Tcells, reduction = 'umap', group.by = "Phase", split.by = "type") +
    scale_color_manual(values=colorblind_vector(3)) 


ggsave(path = "./DataAnalysis/CellType/CD4.Tcells/CellCycle", filename="CellCycle_UMAP_byType.eps", width=10.5, height=3)
```

###Expression Markers in T cells

Like above we also want to look at the expression markers that define clusters and T cell subtypes. First we will make the schex-based reductions. 

```{r}
suppressPackageStartupMessages(library(schex))
Tcells <- make_hexbin(Tcells, 40, dimension_reduction = "UMAP")
```

Here is a list I use in T cell identification, we can run these first and store them in the selected folder.

```{r}
genelist <- c("CD8a", "Cd8b1", "Cd4", "Cd3d", "Foxp3", "Il2ra", "Il7r", "Ccr7", "Ccl4", "Gata3", "Tbx21",  "Cd44", "Cd28", "Sell", "Fas", "Ctla4", "Pdcd1", "Icos", "Havcr2", "Entpd1", "Tigit", "Cd244", "Eomes", "Cd160", "Il10", "Smad3", "Klrg1", "Itga4", "Ifna1", "Ifng", "Cxcr3", "Ccr5", "Il12rb2", "Il18ra", "Il27a", "Stat1", "Stat4", "Ccr3", "Ccr4", "Ccr6", "Ccr10", "Ccr8", "Stat5", "Stat6", "Il4", "Il5", "Il6", "Il9", "Il10", "Il13", "Il17a", "Il17f", "Il21", "Il22", "Tgfbr2", "Ccl20", "Ccl22", "Rorgt", "Rora", "Rorc", "Stat3", "Tnfsf8", "Cxcr5", "Cxcr6", "Maf", "Bcl6", "Gzmb", "Prf1", "Ms4a4b", "Nfat", "Blimp1", "Batf", "Trgc1", "Trgc2", "Trdc", "Il23r", "Top2a", "Birc5", "Cxcl13", "Trac", "Trbc1", "Trbc2", "Cd52", "Nkg7", "Gapdh", "Igfbp4", "Nfkbia", "Ifi27l2a", "Tnfrsf4")
genelist <- toupper(unique(genelist))


dir.create("DataAnalysis/CellType/CD4.Tcells/markers")
dir.create("DataAnalysis/CellType/CD4.Tcells/markers/selected")

DefaultAssay(Tcells) <- "RNA"

for (i in seq_along(genelist)) {
        if (length(which(rownames(Tcells@assays$RNA@counts) == genelist[i])) == 0){
            next() #Need to loop here because plot_hexbin_feature() does not have a built-in function to deal with absence of selected gene
        } else {
        plot <- plot_hexbin_feature(Tcells, feature = genelist[i], type = "counts", action = "prop_0")+ 
             guides(fill=F, color = F) + 
                scale_fill_gradientn(colors = rev(colorblind_vector(13))) 
        ggsave(path = "DataAnalysis/CellType/CD4.Tcells/markers/selected", file = paste0( genelist[i], "_prop.pdf"), plot, height=3, width=3.25)
        }
    }

```

We can also look at the top markers for each cluster by using the *FindAllMarkers()* call.

```{r}
All.markers <- FindAllMarkers(Tcells, assay = "RNA", pseudocount.use = 0.1, only.pos = T) 
write.table(All.markers, file = "DataAnalysis/CellType/CD4.Tcells/markers/FindAllMarkers_output.txt", col.names=NA, sep="\t",append=F)
```

```{r}
All.markers <- read.delim("./DataAnalysis/CellType/CD4.Tcells/markers/FindAllMarkers_output.txt")
top10 <- All.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
top10 <- top10$gene #just want the IDs

dir.create("DataAnalysis/CellType/CD4.Tcells/markers/TopClusterMarkers")
for (i in seq_along(top10)) {
    plot <- plot_hexbin_feature(Tcells, feature = top10[i], type = "counts", action = "prop_0")+ 
             guides(fill=F, color = F) + 
                scale_fill_gradientn(colors = rev(colorblind_vector(13))) 
        ggsave(path = "DataAnalysis/CellType/CD4.Tcells/markers/TopClusterMarkers", file = paste0("Top10markers", "_", top10[i], "_prop.pdf"), plot, height=3, width=3.25)
}
```


```{r}
top10 <- All.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
DotPlot(Tcells, features = unique(top10$gene)) + coord_flip() + 
    scale_color_gradientn(colors = rev(colorblind_vector(11))) + 
    guides(color = F, size = F) +
    scale_size(range = c(0.5,3.5))
ggsave(path = "DataAnalysis/CellType/CD4.Tcells/markers", file = "Top10_dotplot.eps", height=8, width=3)
```


```{r}
genes <- c("CCR7", "LEF1", "SELL", "TCF7", "CD27", "CD28", "CD40LG", "CD44", "CD69", "IL2RA", "TFRC","TNFRSF4", "TNFRSF8", "CCL3", "CCL4","GZMA", "GZMB", "GZMK", "IFNG","NKG7", "PRF1", "CTLA4", "HAVCR2", "LAG3", "PDCD1", "TIGIT")

DotPlot(Tcells, features = genes) + coord_flip() + 
    scale_color_gradientn(colors = rev(colorblind_vector(11))) + 
    guides(color = F, size = F) +
    scale_size(range = c(0.5,3.5))
ggsave(path = "DataAnalysis/CellType/CD4.Tcells/markers", file = "SelectMarkers_dotplot.eps", height=6, width=2.5)
```

***

# Differential Gene Expression

## Comparing tumor vs peripheral blood in just the ccRCC samples

### Overall comparison

```{r}
dir.create("DataAnalysis/CellType/CD4.Tcells/DGE")
TumorOnly <- subset(Tcells, orig.ident == "P1" | orig.ident == "P2" | orig.ident == "P3")
TumorOnly  <- NormalizeData(TumorOnly , assay = "RNA")
TumorOnly <- SetIdent(TumorOnly, value = TumorOnly@meta.data$type)
markers <- FindMarkers(object = TumorOnly, 
                only.pos = F, ident.1 = "T", ident.2="P", min.diff.pct = -Inf, 
                logfc.threshold = -Inf, min.pct = -Inf)
markers$names <- rownames(markers)
markers  <- markers  %>%
        mutate(Difference = pct.1 - pct.2)
mat <- data.frame(markers)
  write.table(mat, file="./DataAnalysis/CellType/CD4.Tcells/DGE/Overall_differentialMarkers_TvPB.txt",
                sep="\t",append=F, row.names = FALSE)
```

### By Cluster comparison
```{r}

DefaultAssay(TumorOnly) <- "RNA"
TumorOnly <- subset(Tcells, orig.ident == "P1" | orig.ident == "P2" | orig.ident == "P3")
clusters <- as.character(unique(TumorOnly@active.ident))
mat <- NULL
for (i in c(1,2,4,5,8)) {
    tmp <- subset(TumorOnly, idents = clusters[i])
    tmp <- NormalizeData(tmp, assay = "RNA") #after subset always normalize RNA
    tmp <- SetIdent(tmp, value = tmp@meta.data$type)
    markers <- FindMarkers(object = tmp, 
                only.pos = F, ident.1 = "T", ident.2="P", min.diff.pct = -Inf, 
                logfc.threshold = -Inf, min.pct = -Inf)
    markers$names <- rownames(markers)
    markers  <- markers  %>%
        mutate(Difference = pct.1 - pct.2)
    
    mat <- data.frame(markers)
    write.table(mat, file=paste("./DataAnalysis/CellType/CD4.Tcells/DGE/TvPB_differentialMarkers_", clusters[i], ".txt", sep=""),
                sep="\t",append=F, row.names = FALSE)
    mat=NULL
}
```

## Comparing tumor vs normal kidney parenchyma

### Overall comparison

```{r}
dir.create("DataAnalysis/CellType/CD4.Tcells/DGE")
subset <- subset(Tcells, Run == "P1_T" |  Run == "P2_T" |  Run == "P3_T" |  Run == "N2_K" |  Run == "N3_K" |  Run == "N4_K")
subset <- NormalizeData(subset  , assay = "RNA")
subset <- SetIdent(subset, value = subset@meta.data$type)
markers <- FindMarkers(object = subset , 
                only.pos = F, ident.1 = "T", ident.2="K", min.diff.pct = -Inf, 
                logfc.threshold = -Inf, min.pct = -Inf)
markers$names <- rownames(markers)
markers  <- markers  %>%
        mutate(Difference = pct.1 - pct.2)
mat <- data.frame(markers)
  write.table(mat, file="./DataAnalysis/CellType/CD4.Tcells/DGE/Overall_differentialMarkers_TvK.txt",
                sep="\t",append=F, row.names = FALSE)
```

### By Cluster comparison
```{r}
subset <- subset(Tcells, Run == "P1_T" |  Run == "P2_T" |  Run == "P3_T" |  Run == "N2_K" |  Run == "N3_K" |  Run == "N4_K")
clusters <- as.character(unique(subset@active.ident))
DefaultAssay(subset) <- "RNA"


mat <- NULL
for (i in seq_along(clusters)) {
    tmp <- subset(subset, idents = clusters[i])
    tmp <- NormalizeData(tmp, assay = "RNA") #after subset always normalize RNA
    tmp <- SetIdent(tmp, value = tmp@meta.data$type)
    markers <- FindMarkers(object = tmp, 
                only.pos = F, ident.1 = "T", ident.2="K", min.diff.pct = -Inf, 
                logfc.threshold = -Inf, min.pct = -Inf)
    markers$names <- rownames(markers)
    markers  <- markers  %>%
        mutate(Difference = pct.1 - pct.2)
    
    mat <- data.frame(markers)
    write.table(mat, file=paste("./DataAnalysis/CellType/CD4.Tcells/DGE/TvK_differentialMarkers_", clusters[i], ".txt", sep=""),
                sep="\t",append=F, row.names = FALSE)
    mat=NULL
}
```

## Comparing peripheral blood of patients to peripheral blood of normal healthy

```{r}
subset <- subset(Tcells, Run == "P1_P" |  Run == "P2_P" |  Run == "P3_P" |  Run == "N1_P")
subset@meta.data$ind <- ifelse(subset@meta.data$Run == "N1_P", "Con", "Tum")
subset <- NormalizeData(subset  , assay = "RNA")
subset <- SetIdent(subset, value = subset@meta.data$ind)
markers <- FindMarkers(object = subset , 
                only.pos = F, ident.1 = "Tum", ident.2="Con", min.diff.pct = -Inf, 
                logfc.threshold = -Inf, min.pct = -Inf)
markers$names <- rownames(markers)
markers  <- markers  %>%
        mutate(Difference = pct.1 - pct.2)
mat <- data.frame(markers)
  write.table(mat, file="./DataAnalysis/CellType/CD4.Tcells/DGE/Overall_differentialMarkers_TumvCon.txt",
                sep="\t",append=F, row.names = FALSE)
```

***

# Trajectory

```{r}
library(slingshot)
dir.create("./DataAnalysis/CellType/CD4.Tcells/slingshot")

sce <- as.SingleCellExperiment(Tcells) 

sds <- slingshot(Embeddings(Tcells, "umap"), clusterLabels = Tcells@active.ident, allow.breaks = TRUE, stretch = 0, reducedDim = "UMAP") #Calcualting the trajectory
```

```{r}
library(scales)
#Making plots more siminmar to ggplot outputs of Seurat
cell_pal <- function(cell_vars, pal_fun,...) {
  if (is.numeric(cell_vars)) {
    pal <- pal_fun(100, ...)
    return(pal[cut(cell_vars, breaks = 100)])
  } else {
    categories <- sort(unique(cell_vars))
    pal <- setNames(pal_fun(length(categories), ...), categories)
    return(pal[cell_vars])
  }
}
#We need color palettes Leiden clusters. These would be the same colors seen in the Seurat plots.

cell_colors_clust <- cell_pal(Tcells@active.ident, hue_pal())

pdf("./DataAnalysis/CellType/CD4.Tcells/slingshot/Trajectory.pdf", height=4, width=4)
plot(reducedDim(sds), col = cell_colors_clust, pch = 16, cex = 0.25)
lines(sds, lwd = 2, type = 'lineages', col = 'black')
dev.off()
pdf("./DataAnalysis/CellType/CD4.Tcells/slingshot/Trajectory2.pdf", height=4, width=4)
plot(reducedDim(sds), col = cell_colors_clust, pch = 16, cex = 0.5)
lines(sds, lwd = 2, col = 'black')
dev.off()
```

```{r}
nc <- 2
pt <- slingPseudotime(sds)
nms <- colnames(pt)
nr <- ceiling(length(nms)/nc)
pal <- rev(colorblind_vector(100))
par(mfrow = c(nr, nc))
pdf("./DataAnalysis/CellType/CD4.Tcells/slingshot/Trajectory2_pseudotime.pdf", height=4, width=4)
for (i in nms) {
  colors <- pal[cut(pt[,i], breaks = 100)]
  plot(reducedDim(sds), col = colors, pch = 16, cex = 0.5, main = i)
  lines(sds, lwd = 2, col = 'black')
}
dev.off()
```

## Using pseudotime to rank genes

```{r}
library(rsample)
library(tidymodels)

Tcells <- FindVariableFeatures(Tcells)

# Get top 498 highly variable genes
top_hvg <- HVFInfo(Tcells) %>% 
  mutate(., bc = rownames(.)) %>% 
  arrange(desc(variance)) %>% 
  top_n(300, variance) %>% 
  pull(bc)
# Prepare data for random forest
dat_use <- t(as.matrix(GetAssayData(Tcells, slot = "data"))[top_hvg,])
dat_use_df <- cbind(slingPseudotime(sds)[,1], dat_use) # Do curve 2, so 2nd columnn
colnames(dat_use_df)[1] <- "pseudotime"
dat_use_df <- as.data.frame(dat_use_df[!is.na(dat_use_df[,1]),])


dat_split <- initial_split(dat_use_df)
dat_train <- training(dat_split)
dat_val <- testing(dat_split)

colnames(dat_train) <- stringr::str_replace(colnames(dat_train), "-", ".") #Replace dash with dots in gene names
colnames(dat_val) <- stringr::str_replace(colnames(dat_val), "-", ".")

#Performing random forest
model <- rand_forest(mtry = 200, trees = 1400, min_n = 15, mode = "regression") %>%
  set_engine("ranger", importance = "impurity", num.threads = 3) %>%
  fit(pseudotime ~ ., data = dat_train) 

#Getting top 200 genes by pseudotime
var_imp <- sort(model$fit$variable.importance, decreasing = TRUE)
top_genes <- names(var_imp)[1:200]
top_genes <- gsub("\\.", "-", top_genes)

output <- data.frame(model$fit$variable.importance)
colnames(output) <- "Var.Imp"
write.table(output, file = "DataAnalysis/CellType/CD4.Tcells/slingshot/VarImp_model.txt", col.names=NA, sep="\t",append=F)
```

```{r}
output <- read.delim("./DataAnalysis/CellType/CD4.Tcells/slingshot/VarImp_model.txt")
subset_out <- output %>% top_n(20, wt = Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD4.Tcells/slingshot/top20_varImp_raw.pdf")

subset_out$Var.Imp <- subset_out$Var.Imp/max(output$Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD4.Tcells/slingshot/top20_varImp_scaled.pdf")
```


### Importance for T-cell-related genes 
```{r}
output <- read.delim("./DataAnalysis/CellType/CD4.Tcells/slingshot/VarImp_model.txt")
genelist <- c("CD8a", "Cd8b1", "Cd4", "Cd3d", "Foxp3", "Il2ra", "Il7r", "Ccr7", "Ccl4", "Gata3", "Tbx21",  "Cd44", "Cd28", "Sell", "Fas", "Ctla4", "Pdcd1", "Icos", "Havcr2", "Entpd1", "Tigit", "Cd244", "Eomes", "Cd160", "Il10", "Smad3", "Klrg1", "Itga4", "Ifna1", "Ifng", "Cxcr3", "Ccr5", "Il12rb2", "Il18ra", "Il27a", "Stat1", "Stat4", "Ccr3", "Ccr4", "Ccr6", "Ccr10", "Ccr8", "Stat5", "Stat6", "Il4", "Il5", "Il6", "Il9", "Il10", "Il13", "Il17a", "Il17f", "Il21", "Il22", "Tgfbr2", "Ccl20", "Ccl22", "Rorgt", "Rora", "Rorc", "Stat3", "Tnfsf8", "Tnfrsf4", "Tnfrsf8","Ccl5", "Cxcr5", "Cxcr6", "Maf", "Bcl6", "Gzmb", "Prf1", "Ms4a4b", "Nfat", "Blimp1", "Batf","Il23r", "Top2a", "Birc5", "Cxcl13", "Nkg7", "Gapdh", "Nfkbia", "Igfbp4", "Cxcl2", "Cxcl14", "Ifngr1", "Cd28", "Cd3e", "Cd3g", "Cd52", "Lag3", "Junb", "Fox", "Ifi27l2a", "Ifitm3", "Gzma")
genelist <- toupper(unique(genelist))

subset_out <- subset(output, X %in% genelist)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD4.Tcells/slingshot/Filtered_varImp_raw.pdf")

subset_out$Var.Imp <- subset_out$Var.Imp/max(output$Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD4.Tcells/slingshot/Filtered_varImp_scaled.pdf")
```

```{r}
PT <- as.data.frame(slingPseudotime(sds))
colnames(PT) <- c("PseudoT1", "PseudoT2" , "PseudoT3")
Tcells <- AddMetaData(Tcells, PT)
saveRDS(Tcells, file = "./data/Processed/CD4.Tcells_FullCluster.rds")
```

***

#Clonotype Analysis

```{r}
library(scRepertoire)
dir.create("./DataAnalysis/CellType/CD4.Tcells/clonotype")
```


```{r}
Tcells@meta.data$cloneType <- factor(Tcells@meta.data$cloneType, levels = c("Hyperexpanded (100 < X <= 500)", "Large (20 < X <= 100)", "Medium (5 < X <= 20)", "Small (1 < X <= 5)", "Single (0 < X <= 1)", NA))
DimPlot(Tcells, group.by = "cloneType") + scale_color_manual(values = c(colorblind_vector(5)), na.value="grey")
ggsave(path = "DataAnalysis/CellType/CD4.Tcells/clonotype", filename="IntegratedObject_byClonotypeFreq.eps", width=6.5, height=3)
```


```{r}
TumorOnly <- subset(Tcells, orig.ident == "P1" | orig.ident ==  "P2" | orig.ident ==  "P3")
CD4 <- expression2List(Tcells, group = "cluster")
TumorOnly  <- expression2List(TumorOnly,  group = "cluster")
```

```{r}
quantContig(CD4, cloneCall="gene+nt", scale = TRUE) + guides(fill = F)
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/UniqueClonotypes.pdf", height=2, width=2.5)

quantContig(TumorOnly, cloneCall="gene+nt", scale = TRUE) + guides(fill = F)
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/UniqueClonotype_TumorOnlys.pdf", height=2, width=2.5)
```

```{r}
compareClonotypes(CD4, numbers = 15, 
                    cloneCall="gene+nt", graph = "alluvial")
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Alluvial_compare.pdf", height=4, width=20)

compareClonotypes(TumorOnly, numbers = 15, 
                    cloneCall="gene+nt", graph = "alluvial")
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Alluvial_compare_TumorOnly.pdf", height=4, width=20)
```

```{r}
clonalHomeostasis(CD4, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Clonal_homeostasis.pdf", height=2, width=5)

clonalProportion(CD4, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Clonal_proportion.pdf", height=2, width=5)

clonalOverlap(CD4, cloneCall = "gene+nt", method = "morisita")
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Clonal_overlap_morisita.pdf", height=3, width=4)

clonalOverlap(CD4, cloneCall = "gene+nt", method = "overlap")
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Clonal_overlap_overlap.pdf", height=3, width=4)

clonalHomeostasis(TumorOnly, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Clonal_homeostasis_TumorOnly.pdf", height=2, width=5)

clonalProportion(TumorOnly, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Clonal_proportion_TumorOnly.pdf", height=2, width=5)

clonalOverlap(TumorOnly, cloneCall = "gene+nt", method = "morisita")
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Clonal_overlap_morisita_TumorOnly.pdf", height=3, width=4)

clonalOverlap(TumorOnly, cloneCall = "gene+nt", method = "overlap")
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Clonal_overlap_overlap_TumorOnly.pdf", height=3, width=4)
```

```{r}
clonalDiversity(CD4, cloneCall = "gene+nt", group = "samples")
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Clonal_Diversity.pdf", height=3, width=4)

clonalDiversity(TumorOnly, cloneCall = "gene+nt", group = "samples")
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Clonal_Diversity_TumorOnly.pdf", height=3, width=4)
```

```{r}
alluvialClonotypes(Tcells, cloneCall = "gene", 
                   y.axes = c("Frequency", "type", "newCluster"), 
                   color = "newCluster") 
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Clonal_Alluvial.pdf", height=3, width=5)

TumorOnly <- subset(Tcells, orig.ident == c("P1", "P2", "P3"))
alluvialClonotypes(TumorOnly, cloneCall = "gene", 
                   y.axes = c("Frequency", "type", "newCluster"), 
                   color = "newCluster") 
ggsave("DataAnalysis/CellType/CD4.Tcells/clonotype/Clonal_Alluvial_TumorOnly.pdf", height=3, width=5)
```

***

#ssGSEA

```{r}
dir.create("./DataAnalysis/CellType/CD4.Tcells/GSEA")
library(escape)
GS <- getGeneSets(library = c("H", "C5"))
ES <- enrichIt(obj = Tcells, gene.sets = GS, groups = 1000, cores = 4)
save(ES, file = "./DataAnalysis/CellType/CD4.Tcells/GSEA/HC5_enrichment.rda")
```


```{r}
load("./DataAnalysis/CellType/CD4.Tcells/GSEA/H_enrichment.rda")
meta <- Tcells[[]]
meta <- merge(meta, ES, by = "row.names")
names <- colnames(ES)
type_sig <- getSignificance(meta[,c("type",names)], group = "type")
cluster_sig <- getSignificance(meta[,c("newCluster",names)], group = "newCluster")
clonotype_sig <- getSignificance(na.omit(meta[,c("cloneType",names)]), group = "cloneType")
colnames(clonotype_sig)[1:5] <- gsub("\\..*", "", colnames(clonotype_sig)[1:5])

write.table(type_sig , file = "DataAnalysis/CellType/CD4.Tcells/GSEA/lmfit_byType.txt", col.names=NA, sep="\t",append=F)
write.table(cluster_sig, file = "DataAnalysis/CellType/CD4.Tcells/GSEA/lmfit_byCluster.txt", col.names=NA, sep="\t",append=F)
write.table(clonotype_sig, file = "DataAnalysis/CellType/CD4.Tcells/GSEA/lmfit_byCloneType.txt", col.names=NA, sep="\t",append=F)
```


```{r}
top10 <- cluster_sig %>% top_n(10, wt=F)
top10 <- rownames(top10)

for (i in seq_along(top10)){
  ridgeEnrichment(meta, gene.set = top10[i], group = "newCluster", add.rug = TRUE) + facet_grid(.~type)
  ggsave(path = "DataAnalysis/CellType/CD4.Tcells/GSEA/", file = paste0("Ridge_ClusterTop10_TypevsCluster_", top10[i], ".pdf"), height=3, width=5)
}

for (i in seq_along(top10)){
  ridgeEnrichment(meta, gene.set = top10[i], group = "newCluster", add.rug = TRUE) 
  ggsave(path = "DataAnalysis/CellType/CD4.Tcells/GSEA/", file = paste0("Ridge_ClusterTop10_Cluster_", top10[i], ".pdf"), height=3, width=5)
}


top10 <- type_sig %>% top_n(10, wt=F)
top10 <- rownames(top10)
for (i in seq_along(top10)){
  ridgeEnrichment(meta, gene.set = top10[i], group = "type", add.rug = TRUE)
  ggsave(path = "DataAnalysis/CellType/CD4.Tcells/GSEA/", file = paste0("Ridge_ClusterTop10_Type_", top10[i], ".pdf"), height=3, width=5)
}


```


