---
title: Integration of multiple experiments for the ccRCC project
author: 
- name: Nick Borcherding
  email: ncborch@gmail.com
  affiliation: Washington University in St. Louis, School of Medicine, St. Louis, MO, USA
date: "June 25, 2020"
output:
  BiocStyle::html_document:
    toc_float: true

---

```{r, echo=FALSE, results="hide", message=FALSE}
knitr::opts_chunk$set(error=FALSE, message=FALSE, warning=FALSE)
library(BiocStyle)
```


# Preperation

## Loading Libaries

In general I like to load libraries here that we will use universally, and then call other libraries when we need them in the code chunks that are relevant. 

```{r}
suppressPackageStartupMessages(library(Seurat))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
```

## Loading Data

```{r}
integrated <- readRDS("./data/Processed/integrated_Cluster.rds")
load("./data/Processed/completeMeta.rda")
```

## Selecting Color Palette

```{r setup, include=FALSE}
colorblind_vector <- colorRampPalette(c("#FF4B20", "#FFB433", "#C6FDEC", "#7AC5FF", "#0348A6"))
```

***

# Initial Subsetting and Analysis

```{r}
Tcells <- subset(integrated, Minor == "CD8.T.Cell")
Tcells <- NormalizeData(Tcells)
rm(integrated)
```

We are also going to create a subdirectory for all T cell analyses:
```{r}
dir.create("./DataAnalysis/CellType/CD8.Tcells")
```


## Reclustering 

Isolating just the T cells, stripping the seurat object of non-"RNA" values using *DietSeurat()* and then spliting the object to create a list that can be fed into **SCTransform()**
```{r eval=FALSE}
Tcells@meta.data$Run <- paste0(Tcells@meta.data$sample, "_", Tcells@meta.data$type) #Define the sequence run
Tcells[["percent.mt"]] <- PercentageFeatureSet(Tcells, pattern = "^MT-") #Calculate %mito

Tcells <- DietSeurat(Tcells, assays = "RNA")
T.list <- SplitObject(object = Tcells , split.by = "sample")

for (i in 1:length(T.list)) {
    T.list[[i]] <- SCTransform(T.list[[i]], verbose = FALSE, conserve.memory=TRUE)
}
rm(Tcells)
```

Integrating the data as above with the whole data set.
```{r eval = FALSE}
options(future.globals.maxSize= 2621440000)
features <- SelectIntegrationFeatures(object.list = T.list, nfeatures = 3000)
T.list <- PrepSCTIntegration(object.list = T.list, anchor.features = features,
verbose = FALSE)
T.anchors <- FindIntegrationAnchors(object.list = T.list, normalization.method = "SCT",
anchor.features = features, verbose = FALSE, k.filter = 50)
Tcells <- IntegrateData(anchorset = T.anchors, normalization.method = "SCT", verbose = FALSE)
rm(T.list)
rm(T.anchors)
saveRDS(Tcells, file = "./data/Processed/CD8.Tcells_Precluster.rds")
```

Now we can process the intgreated T cells and identify clusters.
```{r eval = FALSE}
Tcells <- readRDS("./data/Processed/CD8.Tcells_Precluster.rds")


dir.create("./DataAnalysis/CellType/CD8.Tcells/DimensionInputs")
for (i in c(10,15,20,25,30,35,40)) {
    
    integrated2 <- ScaleData(object = Tcells, verbose = FALSE)
    integrated2 <- RunPCA(object = integrated2, npcs = 40, verbose = FALSE)
    integrated2 <- RunUMAP(object = integrated2, reduction = "pca", 
        dims = 1:i) #dimensions for  UMAP
    for (x in c(10,15,20,25,30,35,40)) { #Dimensions for neighbors
       integrated2 <- FindNeighbors(object = integrated2, dims = 1:x)
       for(y in c(0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) { #Resolution
           integrated2 <- FindClusters(object = integrated2, resolution = y, force.recalc=T)
           plot1 <- DimPlot(integrated2, reduction = "umap", pt.size = 0.5, group.by = "orig.ident")
           ggsave(path = "./DataAnalysis/CellType/CD8.Tcells/DimensionInputs", file = paste("UMAP", i, "_Neighbors", x, "_resolution", y, "_OrigIdent.pdf", sep=""), plot1, width=4.25, height=3)
           plot2 <- DimPlot(object = integrated2, reduction = 'umap', label = T) + NoLegend()
            ggsave(path = "./DataAnalysis/CellType/CD8.Tcells/DimensionInputs", file = paste("UMAP", i, "_Neighbors", x, "_resolution", y, "_clusters.pdf", sep=""), plot2, width=3.5, height=3)
       }
    }
}
rm(integrated2)

```


```{r eval = FALSE}
Tcells <- ScaleData(object = Tcells, verbose = FALSE)
Tcells <- RunPCA(object = Tcells, npcs = 40, verbose = FALSE)
Tcells <- RunUMAP(object = Tcells, reduction = "pca", 
    dims = 1:40)
Tcells <- FindNeighbors(object = Tcells, dims = 1:40, force.recalc = T)
Tcells <- FindClusters(object = Tcells, resolution = 0.8, force.recalc=T)
```


```{r}
DimPlot(object = Tcells, reduction = 'umap', label = T, group.by = "sample") 
DimPlot(object = Tcells, reduction = 'umap', label = T) 
```

##Mnaually editing cluster assignments and removing outliers
```{r}


Tcells <- subset(Tcells, idents = c(0,1,2,3,4,5,6,8,11)) #Cluster 9 and 10 have > 90% NA clonotype

plot <- DimPlot(Tcells, reduction = "umap")
sc4.cells <- CellSelector(plot=plot)
Tcells <- subset(Tcells, cells = sc4.cells, invert = TRUE)

Tcells <- subset(Tcells, idents = 11, invert = T) #Cluster 11 appears to be B cells contamination from healthy peripheral blood
```

```{r}
Tcells@meta.data$newCluster <- Tcells@active.ident
DimPlot(object = Tcells, reduction = 'umap', label = T, ) + NoLegend()
ggsave(path = "./DataAnalysis/CellType/CD8.Tcells", filename="IntegratedObject_byCluster.eps", width=3.5, height=3)
DimPlot(object = Tcells, reduction = 'umap', group.by = "type") 
ggsave(path = "./DataAnalysis/CellType/CD8.Tcells", filename="IntegratedObject_byType.eps", width=4, height=3)

DimPlot(object = Tcells, reduction = 'umap', group.by = "Final_clusters") 
ggsave(path = "./DataAnalysis/CellType/CD8.Tcells", filename="IntegratedObject_byGlobalCluster.eps", width=3.75, height=3)
```



```{r}
freq_table <- table(Tcells@active.ident, Tcells$type)

tTotal <- 12239
pTotal <- 21160
kTotal <- 1778
totals <- c(kTotal,pTotal,tTotal)

for (i in 1:3) { 
    freq_table[,i] <- freq_table[,i]/totals[i]
}

freq_table <- reshape2::melt(freq_table)

ggplot(freq_table, aes(x=as.factor(Var1), y=value, fill=Var2)) + 
  geom_bar(stat="identity", position="fill", color="black", lwd=0.25) + 
  theme(axis.title.x = element_blank()) + 
   scale_fill_manual(values=rev(colorblind_vector(3))) + 
  theme_classic() + 
    xlab("Clusters") + 
    coord_flip() 
ggsave(path = "DataAnalysis/CellType/CD8.Tcells/", file = "relativeContribution_byClusterType_scaled.pdf", height=5, width=5)


ggplot(freq_table, aes(x=as.factor(Var1), y=value, fill=Var2)) + 
  geom_bar(stat="identity", color="black", lwd=0.25) + 
  theme(axis.title.x = element_blank()) + 
   scale_fill_manual(values=rev(colorblind_vector(3))) + 
  theme_classic() + 
    xlab("Clusters") + 
    coord_flip()
ggsave(path = "DataAnalysis/CellType/CD8.Tcells/", file = "totalContribution_byClusterType_scaled.pdf", height=5, width=5)

```

## Cell Cycle 

Cell cycle regression as described in the [Satija lab website](https://satijalab.org/seurat/v3.1/cell_cycle_vignette.html).  

```{r}
cc.genes <- Seurat::cc.genes.updated.2019
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
```

Now we can perform cell cycle scoring with the genes. For now, I am not going to regress using the assingments or save the integrated object with the calculations.

```{r}
DefaultAssay(Tcells) <- "RNA"
Tcells <- CellCycleScoring(Tcells, s.features = s.genes, g2m.features = g2m.genes, set.ident = FALSE)
```

Like above with the contribution by condition (**type**), we can now look at the phases assignments by cluster and condition. We do not need to scale if we seperate **type** into seperate bar graphs.

```{r}
dir.create("./DataAnalysis/CellType/CD8.Tcells/CellCycle")
freq_table <- Tcells[[]]
freq_table <- freq_table[,c("type", "newCluster", "Phase")]
freq_table <- subset(freq_table, Phase != "Undecided") #removing undecided phases
freq_table <- freq_table %>%
    group_by(type, newCluster, Phase) %>%
    summarise(n = n())

freq_table$Phase <- factor(freq_table$Phase, levels = c("G1", "S", "G2M")) #ordering phases

ggplot(freq_table, aes(x=newCluster, y=n, fill=Phase)) + 
  geom_bar(stat="identity", position="fill", color="black", lwd=0.25) + 
  theme(axis.title.x = element_blank())+ 
    facet_grid(type ~.) + 
   scale_fill_manual(values=colorblind_vector(3)) + 
  theme_classic()
ggsave(path = "./DataAnalysis/CellType/CD8.Tcells/CellCycle", file = "CellCycle_byCluster_byType.pdf", height=6, width=6)

ggplot(freq_table, aes(x=newCluster, y=n, fill=Phase)) + 
  stat_summary(geom="bar", position="fill", color="black", lwd=0.25) + 
  theme(axis.title.x = element_blank())+ 
   scale_fill_manual(values=colorblind_vector(3)) + 
  theme_classic()
ggsave(path = "./DataAnalysis/CellType/CD8.Tcells/CellCycle", file = "CellCycle_byCluster.pdf", height=2, width=6)
```


```{r}
Tcells$Phase <- factor(Tcells$Phase, levels = c("G1", "S", "G2M")) #ordering phases
DimPlot(object = Tcells, reduction = 'umap', group.by = "Phase", split.by = "type") +
    scale_color_manual(values=colorblind_vector(3)) 


ggsave(path = "./DataAnalysis/CellType/CD8.Tcells/CellCycle", filename="CellCycle_UMAP_byType.eps", width=10.5, height=3)
```
###Expression Markers in T cells

Like above we also want to look at the expression markers that define clusters and T cell subtypes. First we will make the schex-based reductions. 

```{r}
suppressPackageStartupMessages(library(schex))
Tcells <- make_hexbin(Tcells, 40, dimension_reduction = "UMAP")
```

Here is a list I use in T cell identification, we can run these first and store them in the selected folder.

```{r}
genelist <- c("Cd8a", "Cd8b1", "Cd4", "Cd3d", "Foxp3", "Il2ra", "Il7r", "Ccr7", "Ccl4", "Gata3", "Tbx21",  "Cd44", "Cd28", "Sell", "Fas", "Ctla4", "Pdcd1", "Icos", "Havcr2", "Entpd1", "Tigit", "Cd244", "Eomes", "Cd160", "Il10", "Smad3", "Klrg1", "Itga4", "Ifna1", "Ifng", "Cxcr3", "Ccr5", "Il12rb2", "Il18ra", "Il27a", "Stat1", "Stat4", "Ccr3", "Ccr4", "Ccr6", "Ccr10", "Ccr8", "Stat5", "Stat6", "Il4", "Il5", "Il6", "Il9", "Il10", "Il13", "Il17a", "Il17f", "Il21", "Il22", "Tgfbr2", "Ccl20", "Ccl22", "Rorgt", "Rora", "Rorc", "Stat3", "Tnfsf8", "Cxcr5", "Cxcr6", "Maf", "Bcl6", "Gzmb", "Prf1", "Ms4a4b", "Nfat", "Blimp1", "Batf", "Trgc1", "Trgc2", "Trdc", "Il23r", "Top2a", "Birc5", "Cxcl13", "Trac", "Trbc1", "Trbc2", "Cd52", "Nkg7", "Gapdh", "Igfbp4", "Nfkbia", "Ifi27l2a", "Tnfrsf4", "TCF7")
genelist <- toupper(unique(genelist))


dir.create("DataAnalysis/CellType/CD8.Tcells/markers")
dir.create("DataAnalysis/CellType/CD8.Tcells/markers/selected")

DefaultAssay(Tcells) <- "RNA"

for (i in seq_along(genelist)) {
        if (length(which(rownames(Tcells@assays$RNA@counts) == genelist[i])) == 0){
            next() #Need to loop here because plot_hexbin_feature() does not have a built-in function to deal with absence of selected gene
        } else {
        plot <- plot_hexbin_feature(Tcells, feature = genelist[i], type = "counts", action = "prop_0")+ 
             guides(fill=F, color = F) + 
                scale_fill_gradientn(colors = rev(colorblind_vector(13))) 
        ggsave(path = "DataAnalysis/CellType/CD8.Tcells/markers/selected", file = paste0( genelist[i], "_prop.pdf"), plot, height=3, width=3.25)
        }
    }

```

We can also look at the top markers for each cluster by using the *FindAllMarkers()* call.

```{r}
Tcells  <- NormalizeData(Tcells, assay = "RNA")
All.markers <- FindAllMarkers(Tcells, assay = "RNA", pseudocount.use = 0.1, only.pos = T) 
write.table(All.markers, file = "DataAnalysis/CellType/CD8.Tcells/markers/FindAllMarkers_output.txt", col.names=NA, sep="\t",append=F)
```

```{r}
All.markers <- read.delim("./DataAnalysis/CellType/CD8.Tcells/markers/FindAllMarkers_output.txt")
top10 <- All.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
top10 <- top10$gene #just want the IDs

dir.create("DataAnalysis/CellType/CD8.Tcells/markers/TopClusterMarkers")
for (i in seq_along(top10)) {
    plot <- plot_hexbin_feature(Tcells, feature = top10[i], type = "counts", action = "prop_0")+ 
             guides(fill=F, color = F) + 
                scale_fill_gradientn(colors = rev(colorblind_vector(13))) 
        ggsave(path = "DataAnalysis/CellType/CD8.Tcells/markers/TopClusterMarkers", file = paste0("Top10markers", "_", top10[i], "_prop.pdf"), plot, height=3, width=3.25)
}
```


```{r}
top5 <- All.markers %>% group_by(cluster) %>% top_n(n = 5, wt = avg_logFC)
DotPlot(Tcells, features = unique(top5$gene)) + coord_flip() + 
    scale_color_gradientn(colors = rev(colorblind_vector(11))) + 
    guides(color = F, size = F) +
    scale_size(range = c(0.5,3.5))
ggsave(path = "DataAnalysis/CellType/CD8.Tcells/markers", file = "Top5_dotplot.eps", height=6, width=3.5)
```


```{r}
genes <- c("CCR7", "LEF1", "SELL", "TCF7", "CD27", "CD28", "CD40LG", "CD44", "CD69", "IL2RA", "TFRC","TNFRSF4", "TNFRSF8", "CCL3", "CCL4","GZMA", "GZMB", "GZMK", "IFNG","NKG7", "PRF1", "CTLA4", "HAVCR2", "LAG3", "PDCD1", "TIGIT")

DotPlot(Tcells, features = genes) + coord_flip() + 
    scale_color_gradientn(colors = rev(colorblind_vector(11))) + 
    guides(color = F, size = F) +
    scale_size(range = c(0.5,3.5))
ggsave(path = "DataAnalysis/CellType/CD8.Tcells/markers", file = "SelectMarkers_dotplot.eps", height=6, width=2.5)
```

***

# Differential Gene Expression

## Comparing tumor vs peripheral blood in just the ccRCC samples

### Overall comparison

```{r}
dir.create("DataAnalysis/CellType/CD8.Tcells/DGE")
TumorOnly <- subset(Tcells, orig.ident == "P1" | orig.ident ==  "P2" | orig.ident ==  "P3")
TumorOnly  <- NormalizeData(TumorOnly , assay = "RNA")
TumorOnly <- SetIdent(TumorOnly, value = TumorOnly@meta.data$type)
markers <- FindMarkers(object = TumorOnly, 
                only.pos = F, ident.1 = "T", ident.2="P", min.diff.pct = -Inf, 
                logfc.threshold = -Inf, min.pct = -Inf)
markers$names <- rownames(markers)
markers  <- markers  %>%
        mutate(Difference = pct.1 - pct.2)
mat <- data.frame(markers)
  write.table(mat, file="./DataAnalysis/CellType/CD8.Tcells/DGE/Overall_differentialMarkers_TvPB.txt",
                sep="\t",append=F, row.names = FALSE)
```

### By Cluster comparison
```{r}

DefaultAssay(TumorOnly) <- "RNA"
TumorOnly <- subset(Tcells, orig.ident == "P1" | orig.ident ==  "P2" | orig.ident ==  "P3")
clusters <- as.character(unique(TumorOnly@active.ident))
mat <- NULL
for (i in c(1,2,4:7)) {
    tmp <- subset(TumorOnly, idents = clusters[i])
    tmp <- NormalizeData(tmp, assay = "RNA") #after subset always normalize RNA
    tmp <- SetIdent(tmp, value = tmp@meta.data$type)
    markers <- FindMarkers(object = tmp, 
                only.pos = F, ident.1 = "T", ident.2="P", min.diff.pct = -Inf, 
                logfc.threshold = -Inf, min.pct = -Inf)
    markers$names <- rownames(markers)
    markers  <- markers  %>%
        mutate(Difference = pct.1 - pct.2)
    
    mat <- data.frame(markers)
    write.table(mat, file=paste("./DataAnalysis/CellType/CD8.Tcells/DGE/TvPB_differentialMarkers_", clusters[i], ".txt", sep=""),
                sep="\t",append=F, row.names = FALSE)
    mat=NULL
}
```

## Comparing tumor vs normal kidney parenchyma

### Overall comparison

```{r}
dir.create("DataAnalysis/CellType/CD8.Tcells/DGE")
subset <- subset(Tcells, Run == "P1_T" |  Run == "P2_T" |  Run == "P3_T" |  Run == "N2_K" |  Run == "N3_K" |  Run == "N4_K")
subset <- NormalizeData(subset  , assay = "RNA")
subset <- SetIdent(subset, value = subset@meta.data$type)
markers <- FindMarkers(object = subset , 
                only.pos = F, ident.1 = "T", ident.2="K", min.diff.pct = -Inf, 
                logfc.threshold = -Inf, min.pct = -Inf)
markers$names <- rownames(markers)
markers  <- markers  %>%
        mutate(Difference = pct.1 - pct.2)
mat <- data.frame(markers)
  write.table(mat, file="./DataAnalysis/CellType/CD8.Tcells/DGE/Overall_differentialMarkers_TvK.txt",
                sep="\t",append=F, row.names = FALSE)
```

### By Cluster comparison
```{r}
subset <- subset(Tcells, Run == "P1_T" |  Run == "P2_T" |  Run == "P3_T" |  Run == "N2_K" |  Run == "N3_K" |  Run == "N4_K")
clusters <- as.character(unique(subset@active.ident))
DefaultAssay(subset) <- "RNA"


mat <- NULL
for (i in seq_along(clusters)) {
    tmp <- subset(subset, idents = clusters[i])
    tmp <- NormalizeData(tmp, assay = "RNA") #after subset always normalize RNA
    tmp <- SetIdent(tmp, value = tmp@meta.data$type)
    markers <- FindMarkers(object = tmp, 
                only.pos = F, ident.1 = "T", ident.2="K", min.diff.pct = -Inf, 
                logfc.threshold = -Inf, min.pct = -Inf)
    markers$names <- rownames(markers)
    markers  <- markers  %>%
        mutate(Difference = pct.1 - pct.2)
    
    mat <- data.frame(markers)
    write.table(mat, file=paste("./DataAnalysis/CellType/CD8.Tcells/DGE/TvK_differentialMarkers_", clusters[i], ".txt", sep=""),
                sep="\t",append=F, row.names = FALSE)
    mat=NULL
}
```

## Comparing peripheral blood of patients to peripheral blood of normal healthy

```{r}
subset <- subset(Tcells, Run == "P1_P" |  Run == "P2_P" |  Run == "P3_P" |  Run == "N1_P")
subset@meta.data$ind <- ifelse(subset@meta.data$Run == "N1_P", "Con", "Tum")
subset <- NormalizeData(subset  , assay = "RNA")
subset <- SetIdent(subset, value = subset@meta.data$ind)
markers <- FindMarkers(object = subset , 
                only.pos = F, ident.1 = "Tum", ident.2="Con", min.diff.pct = -Inf, 
                logfc.threshold = -Inf, min.pct = -Inf)
markers$names <- rownames(markers)
markers  <- markers  %>%
        mutate(Difference = pct.1 - pct.2)
mat <- data.frame(markers)
  write.table(mat, file="./DataAnalysis/CellType/CD8.Tcells/DGE/Overall_differentialMarkers_TumvCon.txt",
                sep="\t",append=F, row.names = FALSE)
```

## Visualizing the differential gene expression

```{r}

file_list <- list.files("./DataAnalysis/CellType/CD8.Tcells/DGE/", pattern = ".txt")
files <- file.path(paste0("./DataAnalysis/CellType/CD8.Tcells/DGE/", file_list))

marker_list <- list()
for (i in 1:length(files)) {
    marker_list[[i]] <- read.delim(files[i])
}

prefix <- suffix <- stringr::str_split(file_list, "_", simplify = T)[,1]
suffix <- stringr::str_split(file_list, "_", simplify = T)[,3] #Isolate just the cell type
suffix <- stringr::str_remove(suffix, ".txt") #remove .txt
suffix <- stringr::str_remove(suffix, " ") #remoce spaces
names <- paste0(prefix, "_", suffix)
names(marker_list) <- names #assign the cell type to each element of the list
```

Going to make 2 different plots, the first will take the percent difference from Myocarditis versus healthy cells and place that delta on the x-axis with fold-change on the y-axis. The other graph will be the typical volcano plot. Gene labeling is based first on siginificance and then on the weight of fold-change + 4xDelta Percent. The latter is such that we want to find genes that are not only upregulated by good markers for myocarditis. 

```{r}
library(ggrepel)
dir.create("DataAnalysis/CellType/CD8.Tcells/DGE/visualizations")

for (i in seq_along(marker_list)) {
tmp <- marker_list[[i]]
tmp <- tmp %>%
    mutate(Trend = ifelse(p_val_adj <= 0.05 & avg_logFC > 0, "Up",
                    ifelse(p_val_adj <= 0.05 & avg_logFC < 0, "Down", "None")))
filter <- subset(tmp, p_val_adj <= 0.05 & avg_logFC > 0)
top10 <- filter %>% top_n(n =10, wt = avg_logFC + 4*Difference)

ggplot(tmp, aes(x=Difference, y=avg_logFC)) + 
  geom_point(size=0.5, color="#999999") + 
    geom_point(data=subset(tmp, Trend == "Up" | Trend == "Down"), aes(color = Trend), size=0.75) + 
  theme_classic() + 
    geom_hline(yintercept = 0, lty = 2) + 
    geom_vline(xintercept = 0, lty = 2)  + 
    geom_text_repel(data=subset(tmp, names %in% top10$names), aes(label=names), segment.size = 0.25, size=2.5) + 
    scale_color_manual(values = rev(colorblind_vector(2)))+
    guides(color=F)
ggsave(path = "DataAnalysis/CellType/CD8.Tcells/DGE/visualizations", file = paste0("pctDifference_vs_foldchange_", names(marker_list[i]), ".pdf"), height = 2.5, width=2.5) 

tmp <- tmp %>%
    mutate(p_val_adj = ifelse(p_val_adj == 0, min(p_val_adj), p_val_adj)) #mutate the p-values that are lower than detection

ggplot(tmp, aes(x=avg_logFC, y=-log10(p_val_adj))) + 
    geom_point(size=0.5, color="#999999") + 
    geom_point(data=subset(tmp, Trend == "Up" | Trend == "Down"), aes(color = Trend), size=0.75) + 
    theme_classic() + 
    scale_y_sqrt() +
    geom_vline(xintercept = 0, lty = 2) + 
    geom_hline(yintercept = 1.3, lty = 2) + 
    geom_text_repel(data=subset(tmp, names %in% top10$names), aes(label=names), segment.size = 0.25, size=2.5) + 
    scale_color_manual(values = rev(colorblind_vector(2))) + 
    guides(color = F)
ggsave(path = "DataAnalysis/CellType/CD8.Tcells/DGE/visualizations", file = paste0("VolcanoPlot_", names(marker_list[i]), ".pdf"), height = 2.5, width=2.5)
}
```

***

# Trajectory

```{r}
library(slingshot)
dir.create("./DataAnalysis/CellType/CD8.Tcells/slingshot")

sce <- as.SingleCellExperiment(Tcells) 

sds <- slingshot(Embeddings(Tcells, "umap"), clusterLabels = Tcells@active.ident, allow.breaks = TRUE, stretch = 2, reducedDim = "UMAP", start.clus = 4, end.clus = 0) #Calcualting the trajectory
```

```{r}
library(scales)
#Making plots more siminmar to ggplot outputs of Seurat
cell_pal <- function(cell_vars, pal_fun,...) {
  if (is.numeric(cell_vars)) {
    pal <- pal_fun(100, ...)
    return(pal[cut(cell_vars, breaks = 100)])
  } else {
    categories <- sort(unique(cell_vars))
    pal <- setNames(pal_fun(length(categories), ...), categories)
    return(pal[cell_vars])
  }
}
#We need color palettes Leiden clusters. These would be the same colors seen in the Seurat plots.

cell_colors_clust <- cell_pal(Tcells@active.ident, hue_pal())

pdf("./DataAnalysis/CellType/CD8.Tcells/slingshot/Trajectory.pdf", height=4, width=4)
plot(reducedDim(sds), col = cell_colors_clust, pch = 16, cex = 0.25)
lines(sds, lwd = 2, type = 'lineages', col = 'black')
dev.off()
pdf("./DataAnalysis/CellType/CD8.Tcells/slingshot/Trajectory2.pdf", height=4, width=4)
plot(reducedDim(sds), col = cell_colors_clust, pch = 16, cex = 0.5)
lines(sds, lwd = 2, col = 'black')
dev.off()
```

```{r}
nc <- 2
pt <- slingPseudotime(sds)
nms <- colnames(pt)
nr <- ceiling(length(nms)/nc)
pal <- rev(colorblind_vector(100))
par(mfrow = c(nr, nc))
pdf("./DataAnalysis/CellType/CD8.Tcells/slingshot/Trajectory2_pseudotime.pdf", height=4, width=4)
for (i in nms) {
  colors <- pal[cut(pt[,i], breaks = 100)]
  plot(reducedDim(sds), col = colors, pch = 16, cex = 0.5, main = i)
  lines(sds, lwd = 2, col = 'black')
}
dev.off()
```

## Using pseudotime to rank genes

Branch 3
```{r}
library(rsample)
library(tidymodels)

Tcells <- FindVariableFeatures(Tcells)

# Get top 300 highly variable genes
top_hvg <- HVFInfo(Tcells) %>% 
  mutate(., bc = rownames(.)) %>% 
  arrange(desc(variance)) %>% 
  top_n(300, variance) %>% 
  pull(bc)
# Prepare data for random forest
dat_use <- t(as.matrix(GetAssayData(Tcells, slot = "data"))[top_hvg,])
dat_use_df <- cbind(slingPseudotime(sds)[,2], dat_use) # Do curve 2, so 2nd columnn
colnames(dat_use_df)[1] <- "pseudotime"
dat_use_df <- as.data.frame(dat_use_df[!is.na(dat_use_df[,1]),])
dat_use_df <- janitor::clean_names(dat_use_df)


dat_split <- initial_split(dat_use_df)
dat_train <- training(dat_split)
dat_val <- testing(dat_split)

colnames(dat_train) <- stringr::str_replace(colnames(dat_train), "-", ".") #Replace dash with dots in gene names
colnames(dat_val) <- stringr::str_replace(colnames(dat_val), "-", ".")

dat_train <- na.omit(dat_train)

#Performing random forest
model <- rand_forest(mtry = 200, trees = 1400, min_n = 15, mode = "regression") %>%
  set_engine("ranger", importance = "impurity", num.threads = 3) %>%
  fit(pseudotime ~ ., data = dat_train) 

#Getting top 200 genes by pseudotime
var_imp <- sort(model$fit$variable.importance, decreasing = TRUE)
top_genes <- names(var_imp)[1:200]
top_genes <- gsub("\\.", "-", top_genes)

output <- data.frame(model$fit$variable.importance)
colnames(output) <- "Var.Imp"
write.table(output, file = "DataAnalysis/CellType/CD8.Tcells/slingshot/VarImp_model_B3.txt", col.names=NA, sep="\t",append=F)

output <- read.delim("./DataAnalysis/CellType/CD8.Tcells/slingshot/VarImp_model_B3.txt")
subset_out <- output %>% top_n(20, wt = Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/top20_varImp_raw_B3.pdf")

subset_out$Var.Imp <- subset_out$Var.Imp/max(output$Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/top20_varImp_scaled_B3.pdf")

### Importance for T-cell-related genes 

output <- read.delim("./DataAnalysis/CellType/CD8.Tcells/slingshot/VarImp_model.txt")
genelist <- c("Cd8a", "Cd8b1", "Cd4", "Cd3d", "Foxp3", "Il2ra", "Il7r", "Ccr7", "Ccl4", "Gata3", "Tbx21",  "Cd44", "Cd28", "Sell", "Fas", "Ctla4", "Pdcd1", "Icos", "Havcr2", "Entpd1", "Tigit", "Cd244", "Eomes", "Cd160", "Il10", "Smad3", "Klrg1", "Itga4", "Ifna1", "Ifng", "Cxcr3", "Ccr5", "Il12rb2", "Il18ra", "Il27a", "Stat1", "Stat4", "Ccr3", "Ccr4", "Ccr6", "Ccr10", "Ccr8", "Stat5", "Stat6", "Il4", "Il5", "Il6", "Il9", "Il10", "Il13", "Il17a", "Il17f", "Il21", "Il22", "Tgfbr2", "Ccl20", "Ccl22", "Rorgt", "Rora", "Rorc", "Stat3", "Tnfsf8", "Tnfrsf4", "Tnfrsf8","Ccl5", "Cxcr5", "Cxcr6", "Maf", "Bcl6", "Gzmb", "Prf1", "Ms4a4b", "Nfat", "Blimp1", "Batf","Il23r", "Top2a", "Birc5", "Cxcl13", "Nkg7", "Gapdh", "Nfkbia", "Igfbp4", "Cxcl2", "Cxcl14", "Ifngr1", "Cd28", "Cd3e", "Cd3g", "Cd52", "Lag3", "Junb", "Fox", "Ifi27l2a", "Ifitm3", "Gzma", "TCF7", "LEF1")
genelist <- tolower(unique(genelist))

subset_out <- subset(output, X %in% genelist)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/Filtered_varImp_raw_B3.pdf")

subset_out$Var.Imp <- subset_out$Var.Imp/max(output$Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/Filtered_varImp_scaled_B3.pdf")


##Branch 4

top_hvg <- HVFInfo(Tcells) %>% 
  mutate(., bc = rownames(.)) %>% 
  arrange(desc(variance)) %>% 
  top_n(300, variance) %>% 
  pull(bc)
# Prepare data for random forest
dat_use <- t(as.matrix(GetAssayData(Tcells, slot = "data"))[top_hvg,])
dat_use_df <- cbind(slingPseudotime(sds)[,2], dat_use) # Do curve 2, so 2nd columnn
colnames(dat_use_df)[1] <- "pseudotime"
dat_use_df <- as.data.frame(dat_use_df[!is.na(dat_use_df[,2]),])
dat_use_df <- janitor::clean_names(dat_use_df)


dat_split <- initial_split(dat_use_df)
dat_train <- training(dat_split)
dat_val <- testing(dat_split)

colnames(dat_train) <- stringr::str_replace(colnames(dat_train), "-", ".") #Replace dash with dots in gene names
colnames(dat_val) <- stringr::str_replace(colnames(dat_val), "-", ".")

dat_train <- na.omit(dat_train)
#Performing random forest
model <- rand_forest(mtry = 200, trees = 1400, min_n = 15, mode = "regression") %>%
  set_engine("ranger", importance = "impurity", num.threads = 3) %>%
  fit(pseudotime ~ ., data = dat_train) 

#Getting top 200 genes by pseudotime
var_imp <- sort(model$fit$variable.importance, decreasing = TRUE)
top_genes <- names(var_imp)[1:200]
top_genes <- gsub("\\.", "-", top_genes)

output <- data.frame(model$fit$variable.importance)
colnames(output) <- "Var.Imp"
write.table(output, file = "DataAnalysis/CellType/CD8.Tcells/slingshot/VarImp_model_B4.txt", col.names=NA, sep="\t",append=F)

output <- read.delim("./DataAnalysis/CellType/CD8.Tcells/slingshot/VarImp_model_B4.txt")
subset_out <- output %>% top_n(20, wt = Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/top20_varImp_raw_B4.pdf")

subset_out$Var.Imp <- subset_out$Var.Imp/max(output$Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/top20_varImp_scaled_B4.pdf")

### Importance for T-cell-related genes 

output <- read.delim("./DataAnalysis/CellType/CD8.Tcells/slingshot/VarImp_model_B4.txt")


subset_out <- subset(output, X %in% genelist)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/Filtered_varImp_raw_B4.pdf")

subset_out$Var.Imp <- subset_out$Var.Imp/max(output$Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/Filtered_varImp_scaled_B4.pdf")

##Branch 5

top_hvg <- HVFInfo(Tcells) %>% 
  mutate(., bc = rownames(.)) %>% 
  arrange(desc(variance)) %>% 
  top_n(300, variance) %>% 
  pull(bc)
# Prepare data for random forest
dat_use <- t(as.matrix(GetAssayData(Tcells, slot = "data"))[top_hvg,])
dat_use_df <- cbind(slingPseudotime(sds)[,5], dat_use) # Do curve 2, so 2nd columnn
colnames(dat_use_df)[1] <- "pseudotime"
dat_use_df <- as.data.frame(dat_use_df[!is.na(dat_use_df[,2]),])
dat_use_df <- janitor::clean_names(dat_use_df)


dat_split <- initial_split(dat_use_df)
dat_train <- training(dat_split)
dat_val <- testing(dat_split)

colnames(dat_train) <- stringr::str_replace(colnames(dat_train), "-", ".") #Replace dash with dots in gene names
colnames(dat_val) <- stringr::str_replace(colnames(dat_val), "-", ".")

dat_train <- na.omit(dat_train)
#Performing random forest
model <- rand_forest(mtry = 200, trees = 1400, min_n = 15, mode = "regression") %>%
  set_engine("ranger", importance = "impurity", num.threads = 3) %>%
  fit(pseudotime ~ ., data = dat_train) 

#Getting top 200 genes by pseudotime
var_imp <- sort(model$fit$variable.importance, decreasing = TRUE)
top_genes <- names(var_imp)[1:200]
top_genes <- gsub("\\.", "-", top_genes)

output <- data.frame(model$fit$variable.importance)
colnames(output) <- "Var.Imp"
write.table(output, file = "DataAnalysis/CellType/CD8.Tcells/slingshot/VarImp_model_B5.txt", col.names=NA, sep="\t",append=F)

output <- read.delim("./DataAnalysis/CellType/CD8.Tcells/slingshot/VarImp_model_B5.txt")
subset_out <- output %>% top_n(20, wt = Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/top20_varImp_raw_B5.pdf")

subset_out$Var.Imp <- subset_out$Var.Imp/max(output$Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/top20_varImp_scaled_B5.pdf")

### Importance for T-cell-related genes 

output <- read.delim("./DataAnalysis/CellType/CD8.Tcells/slingshot/VarImp_model_B5.txt")

subset_out <- subset(output, X %in% genelist)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/Filtered_varImp_raw_B5.pdf")

subset_out$Var.Imp <- subset_out$Var.Imp/max(output$Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/Filtered_varImp_scaled_B5.pdf")

##Branch 2

top_hvg <- HVFInfo(Tcells) %>% 
  mutate(., bc = rownames(.)) %>% 
  arrange(desc(variance)) %>% 
  top_n(300, variance) %>% 
  pull(bc)
# Prepare data for random forest
dat_use <- t(as.matrix(GetAssayData(Tcells, slot = "data"))[top_hvg,])
dat_use_df <- cbind(slingPseudotime(sds)[,3], dat_use) # Do curve 2, so 2nd columnn
colnames(dat_use_df)[1] <- "pseudotime"
dat_use_df <- as.data.frame(dat_use_df[!is.na(dat_use_df[,2]),])
dat_use_df <- janitor::clean_names(dat_use_df)


dat_split <- initial_split(dat_use_df)
dat_train <- training(dat_split)
dat_val <- testing(dat_split)

colnames(dat_train) <- stringr::str_replace(colnames(dat_train), "-", ".") #Replace dash with dots in gene names
colnames(dat_val) <- stringr::str_replace(colnames(dat_val), "-", ".")

dat_train <- na.omit(dat_train)
#Performing random forest
model <- rand_forest(mtry = 200, trees = 1400, min_n = 15, mode = "regression") %>%
  set_engine("ranger", importance = "impurity", num.threads = 3) %>%
  fit(pseudotime ~ ., data = dat_train) 

#Getting top 200 genes by pseudotime
var_imp <- sort(model$fit$variable.importance, decreasing = TRUE)
top_genes <- names(var_imp)[1:200]
top_genes <- gsub("\\.", "-", top_genes)

output <- data.frame(model$fit$variable.importance)
colnames(output) <- "Var.Imp"
write.table(output, file = "DataAnalysis/CellType/CD8.Tcells/slingshot/VarImp_model_B2.txt", col.names=NA, sep="\t",append=F)

output <- read.delim("./DataAnalysis/CellType/CD8.Tcells/slingshot/VarImp_model_B2.txt")
subset_out <- output %>% top_n(20, wt = Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/top20_varImp_raw_B2.pdf")

subset_out$Var.Imp <- subset_out$Var.Imp/max(output$Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/top20_varImp_scaled_B2.pdf")

### Importance for T-cell-related genes 

output <- read.delim("./DataAnalysis/CellType/CD8.Tcells/slingshot/VarImp_model_B2.txt")

subset_out <- subset(output, X %in% genelist)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/Filtered_varImp_raw_B2.pdf")

subset_out$Var.Imp <- subset_out$Var.Imp/max(output$Var.Imp)

ggplot(subset_out, aes(x=reorder(X, Var.Imp), y=Var.Imp)) +
    geom_bar(stat = "identity") + 
    theme_classic() + 
    coord_flip() + 
    theme(axis.title.y = element_blank())
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/Filtered_varImp_scaled_B2.pdf")
```


```{r}
library(UpSetR)

files <- list.files("./DataAnalysis/CellType/CD8.Tcells/slingshot", pattern = "txt")[1:4]
names <- stringr::str_remove(files, ".txt") #remove .txt
names <- stringr::str_remove(names, " ") #remoce spaces
names <- as.character(stringr::str_split(names, "_", simplify = T)[,3])
files <- paste0("./DataAnalysis/CellType/CD8.Tcells/slingshot/", files)
df.list <- lapply(files, read.delim)
for (i in seq_along(df.list)) {
    df.list[[i]] <- subset(df.list[[i]], Var.Imp > 100)
    df.list[[i]] <- df.list[[i]][,1]
}
names(df.list) <- names


postscript(file="./DataAnalysis/CellType/CD8.Tcells/slingshot/Overlap_varImp.eps", onefile=FALSE, width=6, height=5)
upset(fromList(df.list), order.by = "freq")
dev.off()
```







```{r}
PT <- as.data.frame(slingPseudotime(sds))
colnames(PT) <- c("PseudoT1", "PseudoT2", "PseudoT3", "PseudoT4", "PseudoT5")
names <- rownames(PT)
PT <- PT %>%
  mutate(PT1 = ifelse(is.na(PseudoT1), "No", "Yes")) %>%
  mutate(PT2 = ifelse(is.na(PseudoT2), "No", "Yes")) %>%
  mutate(PT3 = ifelse(is.na(PseudoT3), "No", "Yes")) %>%
  mutate(PT4 = ifelse(is.na(PseudoT4), "No", "Yes")) %>%
  mutate(PT5 = ifelse(is.na(PseudoT5), "No", "Yes"))
rownames(PT) <- names

Tcells <- AddMetaData(Tcells, PT)
saveRDS(Tcells, file = "./data/Processed/CD8.Tcells_FullCluster.rds")
Tcells <- readRDS("./data/Processed/CD8.Tcells_FullCluster.rds")
```




```{r}
meta <- Tcells[[]]

ggplot(meta, aes(x=type, y = PseudoT1)) +
  geom_boxplot()
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/PT1_timeBytype.pdf", height=2, width=3)

ggplot(meta, aes(x=type, y = PseudoT2)) +
  geom_boxplot()
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/PT2_timeBytype.pdf", height=2, width=3)

ggplot(meta, aes(x=type, y = PseudoT3)) +
  geom_boxplot()
ggsave("./DataAnalysis/CellType/CD8.Tcells/slingshot/PT3_timeBytype.pdf", height=2, width=3)
```


***

#Clonotype Analysis

```{r}
library(scRepertoire)
dir.create("./DataAnalysis/CellType/CD8.Tcells/clonotype")
```


```{r}
Tcells@meta.data$cloneType <- factor(Tcells@meta.data$cloneType, levels = c("Hyperexpanded (100 < X <= 500)", "Large (20 < X <= 100)", "Medium (5 < X <= 20)", "Small (1 < X <= 5)", "Single (0 < X <= 1)", NA))
DimPlot(Tcells, group.by = "cloneType") + scale_color_manual(values = c(colorblind_vector(5)), na.value="grey")
ggsave(path = "DataAnalysis/CellType/CD8.Tcells/clonotype", filename="IntegratedObject_byClonotypeFreq.eps", width=6.5, height=3)
```


```{r}
TumorOnly <- subset(Tcells, orig.ident == "P1" | orig.ident ==  "P2" | orig.ident ==  "P3")
CD8 <- expression2List(Tcells, group = "cluster")
Run  <- expression2List(TumorOnly,  group = "Run")
integrated  <- expression2List(TumorOnly,  group = "Final_clusters")
TumorOnly  <- expression2List(TumorOnly,  group = "cluster")

P1 <- subset(Tcells, orig.ident == "P1")
P1 <- expression2List(P1,  group = "cluster")

P2 <- subset(Tcells, orig.ident == "P2")
P2 <- expression2List(P2,  group = "cluster")

P3 <- subset(Tcells, orig.ident == "P3")
P3 <- expression2List(P3,  group = "cluster")

clonalProportion(P1, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_proportion_P1.pdf", height=2, width=5)

clonalProportion(P2, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_proportion_P2.pdf", height=2, width=5)

clonalProportion(P3, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_proportion_P3.pdf", height=2, width=5)

compareClonotypes(P1, numbers = 15, 
                    cloneCall="gene+nt", graph = "alluvial")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Alluvial_compare_P1.pdf", height=4, width=20)

compareClonotypes(P2, numbers = 15, 
                    cloneCall="gene+nt", graph = "alluvial")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Alluvial_compare_P2.pdf", height=4, width=20)

compareClonotypes(P3, numbers = 15, 
                    cloneCall="gene+nt", graph = "alluvial")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Alluvial_compare_P3.pdf", height=4, width=20)
```


```{r}
quantContig(CD8, cloneCall="gene+nt", scale = TRUE) + guides(fill = F)
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/UniqueClonotypes.pdf", height=2, width=2.5)

quantContig(TumorOnly, cloneCall="gene+nt", scale = TRUE) + guides(fill = F)
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/UniqueClonotype_TumorOnlys.pdf", height=2, width=2.5)
```

```{r}
compareClonotypes(CD8, numbers = 15, 
                    cloneCall="gene+nt", graph = "alluvial")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Alluvial_compare.pdf", height=4, width=20)

compareClonotypes(TumorOnly, numbers = 15, 
                    cloneCall="gene+nt", graph = "alluvial")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Alluvial_compare_TumorOnly.pdf", height=4, width=20)
```

```{r}
clonalHomeostasis(CD8, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_homeostasis.pdf", height=2, width=5)

clonalProportion(CD8, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_proportion.pdf", height=2, width=5)

clonalOverlap(CD8, cloneCall = "gene+nt", method = "morisita")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_overlap_morisita.pdf", height=4, width=5)

clonalOverlap(CD8, cloneCall = "gene+nt", method = "overlap")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_overlap_overlap.pdf", height=4, width=5)

clonalHomeostasis(TumorOnly, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_homeostasis_TumorOnly.pdf", height=2, width=5)

clonalProportion(TumorOnly, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_proportion_TumorOnly.pdf", height=2, width=5)

clonalOverlap(TumorOnly, cloneCall = "gene+nt", method = "morisita") + 
    scale_fill_gradientn(colors = rev(colorblind_vector(11)), na.value = "white")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_overlap_morisita_TumorOnly.pdf", height=4, width=5)

clonalOverlap(TumorOnly, cloneCall = "gene+nt", method = "overlap") + 
    scale_fill_gradientn(colors = rev(colorblind_vector(11)), na.value = "white")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_overlap_overlap_TumorOnly.pdf", height=4, width=5)
```

```{r}
clonalDiversity(CD8, cloneCall = "gene+nt", group = "samples")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_Diversity.pdf", height=3, width=4)

clonalDiversity(TumorOnly, cloneCall = "gene+nt", group = "samples")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_Diversity_TumorOnly.pdf", height=3, width=4)
```

```{r}
alluvialClonotypes(Tcells, cloneCall = "gene", 
                   y.axes = c("Frequency", "type", "newCluster"), 
                   color = "newCluster") 
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_Alluvial.pdf", height=3, width=5)

TumorOnly <- subset(Tcells, orig.ident == "P1" | orig.ident ==  "P2" | orig.ident ==  "P3")
alluvialClonotypes(TumorOnly, cloneCall = "gene", 
                   y.axes = c("Frequency", "type", "newCluster"), 
                   color = "newCluster") 
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_Alluvial_TumorOnly.pdf", height=3, width=5)
```


```{r}
alluvialClonotypes(TumorOnly, cloneCall = "gene+nt", 
                   y.axes = c("Frequency", "type", "Final_clusters", "sample"), 
                   color = c("TRAV12-2.TRAJ42.TRAC_TGTGCCGTGAACGGGGGAAGCCAAGGAAATCTCATCTTT_TRBV20-1.TRBJ2-3.TRBD2.TRBC2_TGCAGTGCTGAGCGGGAAGACACAGATACGCAGTATTTT", "NA_NA_TRBV24-1.TRBJ1-4.TRBD1.TRBC1_TGTGCCACCAGTGCGACCCTCAGGGTGGTGGCAGAAAAACTGTTTTTT")) 
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/TumorOnly_Clonal_Alluvial_highlight.pdf", height=3, width=5)
```

Pseudotime Assignments


```{r}
meta <- Tcells[[]]
unique <- stringr::str_sort(as.character(unique(meta[,"PT1"])), numeric = TRUE)
    df <- NULL
    for (i in seq_along(unique)) {
        subset <- subset(meta, meta[,"PT1"] == unique[i])
        df[[i]] <- subset
    }
    names(df) <- unique
    
quantContig(df, cloneCall="gene+nt", scale = T) + guides(fill = F)
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/UniqueClonotypes_Pseudotime1.pdf", height=2, width=2.5)

clonalHomeostasis(df, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_homeostasis_Pseudotime1.pdf", height=2, width=5)

clonalProportion(df, cloneCall = "gene+nt", split = c(5, 20, 100, 1000, 10000, 30000))
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_proportion_Pseudotime1.pdf", height=2, width=5)

unique <- stringr::str_sort(as.character(unique(meta[,"PT2"])), numeric = TRUE)
    df <- NULL
    for (i in seq_along(unique)) {
        subset <- subset(meta, meta[,"PT2"] == unique[i])
        df[[i]] <- subset
    }
    names(df) <- unique
    
quantContig(df, cloneCall="gene+nt", scale = T) + guides(fill = F)
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/UniqueClonotypes_Pseudotime2.pdf", height=2, width=2.5)

clonalHomeostasis(df, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_homeostasis_Pseudotime2.pdf", height=2, width=5)

clonalProportion(df, cloneCall = "gene+nt", split = c(5, 20, 100, 1000, 10000, 30000))
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_proportion_Pseudotime2.pdf", height=2, width=5)


unique <- stringr::str_sort(as.character(unique(meta[,"PT3"])), numeric = TRUE)
    df <- NULL
    for (i in seq_along(unique)) {
        subset <- subset(meta, meta[,"PT3"] == unique[i])
        df[[i]] <- subset
    }
    names(df) <- unique
    
quantContig(df, cloneCall="gene+nt", scale = T) + guides(fill = F)
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/UniqueClonotypes_Pseudotime3.pdf", height=2, width=2.5)

clonalHomeostasis(df, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_homeostasis_Pseudotime3.pdf", height=2, width=5)

clonalProportion(df, cloneCall = "gene+nt", split = c(5, 20, 100, 1000, 10000, 30000))
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_proportion_Pseudotime3.pdf", height=2, width=5)


unique <- stringr::str_sort(as.character(unique(meta[,"PT4"])), numeric = TRUE)
    df <- NULL
    for (i in seq_along(unique)) {
        subset <- subset(meta, meta[,"PT4"] == unique[i])
        df[[i]] <- subset
    }
    names(df) <- unique
    
quantContig(df, cloneCall="gene+nt", scale = T) + guides(fill = F)
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/UniqueClonotypes_Pseudotime4.pdf", height=2, width=2.5)

clonalHomeostasis(df, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_homeostasis_Pseudotime4.pdf", height=2, width=5)

clonalProportion(df, cloneCall = "gene+nt", split = c(5, 20, 100, 1000, 10000, 30000))
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_proportion_Pseudotime4.pdf", height=2, width=5)

unique <- stringr::str_sort(as.character(unique(meta[,"PT5"])), numeric = TRUE)
    df <- NULL
    for (i in seq_along(unique)) {
        subset <- subset(meta, meta[,"PT5"] == unique[i])
        df[[i]] <- subset
    }
    names(df) <- unique
    
quantContig(df, cloneCall="gene+nt", scale = T) + guides(fill = F)
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/UniqueClonotypes_Pseudotime5.pdf", height=2, width=2.5)

clonalHomeostasis(df, cloneCall = "gene+nt")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_homeostasis_Pseudotime5.pdf", height=2, width=5)

clonalProportion(df, cloneCall = "gene+nt", split = c(5, 20, 100, 1000, 10000, 30000))
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_proportion_Pseudotime5.pdf", height=2, width=5)
```

```{r}
alluvialClonotypes(Tcells, cloneCall = "gene", 
                   y.axes = c("Frequency", "newCluster", "PT1", "PT2", "PT3"), 
                   color = "newCluster") 
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_Alluvial_AllPseudotime.pdf", height=3, width=7)

alluvialClonotypes(Tcells, cloneCall = "gene", 
                   y.axes = c("Frequency", "Run", "PT1", "PT2", "PT3"), 
                   color = "Run") 
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Clonal_Alluvial_AllPseudotime_byRun.pdf", height=3, width=7)
```

```{r}
quantContig(Run, cloneCall = "gene+nt", scale = T)
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/PtUnique_TumorOnly.pdf", height=3, width=5)

clonalProportion(Run, cloneCall = "gene+nt")

ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/PtProportion_TumorOnly.pdf", height=3, width=5)

compareClonotypes(Run, samples = c("P1_P", "P1_T"), numbers = 10)
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Pt1_alluvialCompare_TumorOnly.pdf", height=3, width=20)

compareClonotypes(Run, samples = c("P2_P", "P2_T"), numbers = 10)
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Pt2_alluvialCompare_TumorOnly.pdf", height=3, width=20)

compareClonotypes(Run, samples = c("P3_P", "P3_T"), numbers = 10)
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/Pt3_alluvialCompare_TumorOnly.pdf", height=3, width=20)

clonalOverlap(Run, cloneCall = "gene+nt") + 
    scale_fill_gradientn(colors = rev(colorblind_vector(11)), na.value = "white")
ggsave("DataAnalysis/CellType/CD8.Tcells/clonotype/PT_OVERLAP.pdf", height=4, width = 5)
```

```{r}
clonalOverlap <- function(df, cloneCall = c("gene", "nt", "aa", "gene+nt"), 
                                method = c("overlap", "morisita"), 
                                exportTable = FALSE){
    cloneCall <- theCall(cloneCall)
    df <- df[order(names(df))]
    values <- str_sort(as.character(unique(names(df))), numeric = TRUE)
    df <- df[quiet(dput(values))]
    num_samples <- length(df[])
    names_samples <- names(df)
    coef_matrix <- data.frame(matrix(NA, num_samples, num_samples))
    colnames(coef_matrix) <- names_samples
    rownames(coef_matrix) <- names_samples
    length <- seq_len(num_samples)
    if (method == "overlap") {
        coef_matrix <- overlapIndex(df, length, cloneCall, coef_matrix)
    } else if (method == "morisita") {
        coef_matrix <- morisitaIndex(df, length, cloneCall, coef_matrix)}
    coef_matrix$names <- rownames(coef_matrix)
    if (exportTable == TRUE) { return(coef_matrix) }
    coef_matrix <- suppressMessages(melt(coef_matrix))[,-1]
    col <- colorblind_vector(7)
    coef_matrix$variable <- factor(coef_matrix$variable, levels = values)
    coef_matrix$names <- factor(coef_matrix$names, levels = values)
    plot <- ggplot(coef_matrix, aes(x=names, y=variable, fill=value)) +
            geom_tile() + labs(fill = method) +
            geom_text(aes(label = round(value, digits = 3))) +
            scale_fill_gradientn(colors = colorblind_vector(11), na.value = "white") +
            theme_classic() + theme(axis.title = element_blank())
}


```


***

#ssGSEA

```{r}
dir.create("./DataAnalysis/CellType/CD8.Tcells/GSEA")
#library(escape)
GS <- getGeneSets(library = c("H", "C5"))
ES <- enrichIt(obj = Tcells, gene.sets = GS, groups = 1000, cores = 4)
save(ES, file = "./DataAnalysis/CellType/CD8.Tcells/GSEA/H_C5_enrichment.rda")
```


```{r}
load("./DataAnalysis/CellType/CD8.Tcells/GSEA/H_C5_enrichment.rda")
meta <- Tcells[[]]
meta <- merge(meta, ES, by = "row.names")
names <- colnames(ES)
type_sig <- getSignificance(meta[,c("type",names)], group = "type")
cluster_sig <- getSignificance(meta[,c("newCluster",names)], group = "newCluster")
clonotype_sig <- getSignificance(na.omit(meta[,c("cloneType",names)]), group = "cloneType")
colnames(clonotype_sig)[1:5] <- gsub("\\..*", "", colnames(clonotype_sig)[1:5])

write.table(type_sig , file = "DataAnalysis/CellType/CD8.Tcells/GSEA/lmfit_byType.txt", col.names=NA, sep="\t",append=F)
write.table(cluster_sig, file = "DataAnalysis/CellType/CD8.Tcells/GSEA/lmfit_byCluster.txt", col.names=NA, sep="\t",append=F)
write.table(clonotype_sig, file = "DataAnalysis/CellType/CD8.Tcells/GSEA/lmfit_byCloneType.txt", col.names=NA, sep="\t",append=F)
```

```{r}
getSignificance <- function(enriched, group = NULL, fit = "linear.model") {
    group2 <- enriched[,group]
    gr_names <- unique(group2)
    input <- select_if(enriched, is.numeric)
    output <- NULL
    if (fit == "linear.model") {
        group2 <- factor(group2)
        design <- model.matrix(~ -1 + group2)
        colnames(design) <- levels(group2)
        fit <- lmFit(t(input), design)
        fit <- eBayes(fit)
        output <- topTable(fit, number = ncol(input))
    } else if (fit == "T.test") {
        if (length(unique(group2)) != 2) {
            stop("Ensure the group selection has only two levels for T.test 
                fit") 
        } else {
        out <- lapply(input, function(x) t.test(x ~ group2))
        for (i in seq_along(out)) {
            df <- out[[i]]
            mat <- c(df$statistic, df$conf.int[[1]], df$conf.int[[2]],
                        df$estimate[1], df$estimate[2], df$p.value)
            output <- rbind(output,mat)
        }
        output <- as.data.frame(output)
        colnames(output) <- c("T_statistic", "CI_low", "CI_high", 
            paste0("mean_", gr_names[1]), paste0("mean_", gr_names[2]), 
            "p.value")
        rownames(output) <- colnames(input)
        output$FDR <- p.adjust(output$p.value) }
    } else if (fit == "ANOVA") {
        if (unique(group2) <= 2) {
            stop("Ensure the group selection has more than two levels 
                for ANOVA fit") }
        out <- lapply(input, function(x) aov(x ~ group2))
        for (i in seq_along(out)) {
            df <- out[[i]]
            fval <- summary(df)[[1]]$'F value'[[1]]
            pval <- summary(df)[[1]]$'Pr(>F)'[[1]]
            output <- rbind(output,mat)
        }
        output <- as.data.frame(output)
        colnames(output) <- c("f.value", "p.value")
        rownames(output) <- colnames(input)
        output$FDR <- p.adjust(output$p.value)
    }
    return(output)
}
```

```{r}
PT1_sig <- getSignificance(meta[,c("PT1",names)], group = "PT1", fit = "T.test")
write.table(PT1_sig , file = "DataAnalysis/CellType/CD8.Tcells/GSEA/lmfit_byPT1.txt", col.names=NA, sep="\t",append=F)
PT2_sig <- getSignificance(meta[,c("PT2",names)], group = "PT2", fit = "T.test")
write.table(PT2_sig , file = "DataAnalysis/CellType/CD8.Tcells/GSEA/lmfit_byPT2.txt", col.names=NA, sep="\t",append=F)
PT3_sig <- getSignificance(meta[,c("PT3",names)], group = "PT3", fit = "T.test")
write.table(PT3_sig , file = "DataAnalysis/CellType/CD8.Tcells/GSEA/lmfit_byPT3.txt", col.names=NA, sep="\t",append=F)
```

```{r}
x <- psych::corr.test(meta[,names], meta$PseudoT1, method = "spearman")
x <- data.frame(x$r, x$p, x$se, x$t, x$n)
colnames(x) <- c("rho", "Pval", "SE", "Tval", "n")
x$FDR <- p.adjust(x$Pval)
write.table(x , file = "DataAnalysis/CellType/CD8.Tcells/GSEA/Corr_PT1.txt", col.names=NA, sep="\t",append=F)

x <- psych::corr.test(meta[,names], meta$PseudoT2, method = "spearman")
x <- data.frame(x$r, x$p, x$se, x$t, x$n)
colnames(x) <- c("rho", "Pval", "SE", "Tval", "n")
x$FDR <- p.adjust(x$Pval)
write.table(x , file = "DataAnalysis/CellType/CD8.Tcells/GSEA/Corr_PT2.txt", col.names=NA, sep="\t",append=F)

x <- psych::corr.test(meta[,names], meta$PseudoT3, method = "spearman")
x <- data.frame(x$r, x$p, x$se, x$t, x$n)
colnames(x) <- c("rho", "Pval", "SE", "Tval", "n")
x$FDR <- p.adjust(x$Pval)
write.table(x , file = "DataAnalysis/CellType/CD8.Tcells/GSEA/Corr_PT3.txt", col.names=NA, sep="\t",append=F)
```




***

Russ requested pathways

```{r}
sign <- read.delim("./data/immune_sign.txt")[,1:20]
sign <- as.list(sign)
melanoma <- read.delim("./data/melanoma_GS.txt")
melanoma <- as.list(melanoma)
```

```{r}

full <- c(sign, melanoma)
unique <- names(full) 
list <- list()
for (i in seq_along(unique)) {
    tmp <- full[[i]]
    tmp <- tmp[tmp != ""]
    tmp <- unique(toupper(tmp))
    tmp <- GSEABase::GeneSet(tmp, setName=paste(unique[i]))
    list[[i]] <- tmp
    
}

list <- GSEABase::GeneSetCollection(list)
```


```{r}
ES2 <- enrichIt(obj = Tcells, gene.sets = list, groups = 1000, cores = 4)
save(ES2, file = "./DataAnalysis/CellType/CD8.Tcells/GSEA/Requested_enrichment.rda")
```

```{r}
Tcells <- AddMetaData(Tcells, ES2)
```

```{r}
for (i in seq_along(unique)) {
  plot <- FeaturePlot(Tcells, features = unique[i]) + 
  scale_color_gradientn(colors = rev(colorblind_vector(11)))
  plot <- plot + NoLegend()
  ggsave(path = "./DataAnalysis/CellType/CD8.Tcells/GSEA", filename=paste0(unique[i], ".eps"), plot, width=3.5, height=3)
}
```

```{r}
ES2 <- get(load("./DataAnalysis/CellType/CD8.Tcells/GSEA/Requested_enrichment.rda"))
```

```{r}
filter <- ES[,grepl("HALLMARK",colnames(ES))]
filter <- cbind(filter,ES2)
meta <- Tcells[[]]
meta <- merge(meta, filter, by = "row.names")
```

```{r}
heatmap <- meta[, c("newCluster", colnames(filter))]
melted <- reshape2::melt(heatmap, id.vars = c("newCluster"))
meanvalues <- melted %>%
  group_by(newCluster, variable) %>%
  summarise(median(value))

matrix <- reshape2::dcast(meanvalues, newCluster ~ variable)
rownames(matrix) <- matrix[,1]
matrix <- matrix[,-1]

pdf("test.pdf", height=12, width=5)
pheatmap::pheatmap(t(matrix), color = rev(colorblind_vector(50)), scale = "row", fontsize_row = 3, cluster_rows = T, cluster_cols = T)
dev.off()

GOI <- c("HALLMARK_INFLAMMATORY_RESPONSE", "T1_Interferon", "T2_Interferon", "Activated", "Exhuasted", "T_Cell_Terminal_Differentiation", "HALLMARK_HYPOXIA", "HALLMARK_IL2_STAT5_SIGNALING","HALLMARK_IL6_JAK_STAT3_SIGNALING","HALLMARK_NOTCH_SIGNALING", "HALLMARK_TGF_BETA_SIGNALING","HALLMARK_HEDGEHOG_SIGNALING","HALLMARK_DNA_REPAIR", "Cytolytic", "Glycolysis", "TCA_cycle")
sub <- matrix[,colnames(matrix) %in% GOI]

pdf("./DataAnalysis/CellType/CD8.Tcells/GSEA/SelectGSEAheatmap.pdf", height=4, width=4)
pheatmap::pheatmap(t(sub), color = rev(colorblind_vector(50)), scale = "row", fontsize_row = 3, cluster_rows = T, cluster_cols = F)
dev.off()
```

```{r}
library(pathwork)

branches <- c("PT1", "PT2", "PT3", "PT4", "PT5")
sets <- colnames(filter)
for(j in seq_along(sets)) {
    plot_list <- list()
    for (i in seq_along(branches)) {
        plot_list[[i]] <- ggplot(meta, aes(x=meta[,branches[i]], y=meta[,sets[j]])) + 
                            geom_boxplot(outlier.alpha = 0) +
                            theme_classic() + 
                            ylab(sets[j]) + 
                            theme(axis.title.x = element_blank(),
                                    axis.ticks.x = element_blank(), 
                                    axis.text.x = element_blank())
    }
}
plot_list
```

```{r}
library(patchwork)
r1 <- ggplot(meta, aes(PseudoT1, y=GD8_G)) + geom_smooth(method = "loess", se=F) + theme_classic() +coord_cartesian(ylim= c(0.2,0.7))
r2 <- ggplot(meta, aes(PseudoT2, y=GD8_G)) + geom_smooth(method = "loess", se=F) + theme_classic()+coord_cartesian(ylim=  c(0.2,0.7))
r3 <- ggplot(meta, aes(PseudoT3, y=GD8_G)) + geom_smooth(method = "loess", se=F) + theme_classic()+coord_cartesian(ylim = c(0.2,0.7))
r4 <- ggplot(meta, aes(PseudoT4, y=GD8_G)) + geom_smooth(method = "loess", se=F) + theme_classic()+coord_cartesian(ylim=  c(0.2,0.7))
r5 <- ggplot(meta, aes(PseudoT5, y=GD8_G)) + geom_smooth(method = "loess", se=F) + theme_classic()+coord_cartesian(ylim= c(0.2,0.7))


r4/r3/r1/r2/r5
ggsave("./DataAnalysis/CellType/CD8.Tcells/GSEA/responders.pdf", height=4,width=5)

nr1 <- ggplot(meta, aes(PseudoT1, y=CD8_B)) + geom_smooth(method = "loess", se=F) + theme_classic() +coord_cartesian(ylim= c(0.4,0.6))
nr2 <- ggplot(meta, aes(PseudoT2, y=CD8_B)) + geom_smooth(method = "loess", se=F) + theme_classic()+coord_cartesian(ylim= c(0.4,0.6))
nr3 <- ggplot(meta, aes(PseudoT3, y=CD8_B)) + geom_smooth(method = "loess", se=F) + theme_classic()+coord_cartesian(ylim= c(0.4,0.6))
nr4 <- ggplot(meta, aes(PseudoT4, y=CD8_B)) + geom_smooth(method = "loess", se=F) + theme_classic()+coord_cartesian(ylim= c(0.4,0.6))
nr5 <- ggplot(meta, aes(PseudoT5, y=CD8_B)) + geom_smooth(method = "loess", se=F) + theme_classic()+coord_cartesian(ylim= c(0.4,0.6))

nr4/nr3/nr1/nr2/nr5
ggsave("./DataAnalysis/CellType/CD8.Tcells/GSEA/nonresponders.pdf", height=4,width=5)
```

***

```{r}
Tcells <- NormalizeData(Tcells)
markers <- FindAllMarkers(Tcells, logfc = 0.0, min.pct = 0, pseudocount.use = 0.1, only.pos = TRUE)
write.table(markers, file="./DataAnalysis/CellType/CD8.Tcells/FindAllMarkers.txt",
                sep="\t",append=F, row.names = FALSE)
```

